Vagueness, Noise, and Communication

### 2014-09-18 paper/03_simulations.tex
more on diffusion; Nosofsky similarity plots

@@

>>>>

Confusion of states should intuitively be a function of their
perceptual similarity. Here and in the following, we assume that the
state space of the sim-max game consists of $n \ge 2$ states that are
equally spaced across the unit interval, including $0$ and $1$. We
assume here that the distance $\card{\mystate{i} - \mystate{j}}$ is
the objective, physical similarity between two states $\mystate{i}$
and $\mystate{j}$. Physical distance space feeds into a perceptual
similarity function, as described by
\citet{Nosofsky1986:Attention-Simil}:
\begin{align}
  \label{eq:NosofskySim}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \alpha) = \expo \left ( -
    \frac{\card{\mystate{i} - \mystate{j}}^s}{ \alpha^2} \right ) \,,
\end{align}
where $\alpha$ is an imprecision parameter. When $\alpha=0$ agents
perfectly discriminate between states; when $\alpha \rightarrow
\infty$ agents cannot discriminate states at
all. Figure~\ref{fig:NosofskySim} gives an impression of
Nosofsky-similarity for different parameter values. Other
formalizations of perceptual similarity are possible, including ones
that allow for different discriminability in different areas of the
state space, but we stick with Nosofsky's similarity function for the
time being, because it is a mathematically simple, yet established
notion in mathematical psychology.

  \includegraphics[width=\textwidth]{plots/NosofskySim.pdf}

  \caption{Examples for Nosofsky-similarity (non-normalized) and
    derived probability of state confusion (normalized) for different
    values of $\alpha$ and $\card{\States} = 50$.}
  \label{fig:NosofskySim}
\end{figure}

We further assume that the probability of confusing any two states
$\mystate{i}$ and $\mystate{j}$ is proportional to their perceived
similarity and therefore obtained by normalization:
\begin{align*}
  C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \alpha)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \alpha)} \,.
\end{align*}
Confusion of states is then also a function of imprecision parameter
$\alpha$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\alpha
= 0$ the confusion matrix has $C_{ii} = 1$; for $\alpha > 0$ $C$ is
positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\alpha
\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$ for all
$i$ and $j$.

@@

\newpage
